Figure two

### 2025-12-05 draftFiguresPaper.Rmd
fix bug year survey item

# 2) get total surveys by country
surveys_by_country <- surveyCountries_long_clean %>%
  filter(Inclusion == 1) %>% 
  group_by(iso3, .drop = F) %>% 
  count(iso3, name = "n_surveys") %>%
  arrange(desc(n_surveys))

@@

df_cat_long <- surveyCountries_long_clean %>%
  pivot_longer(
    cols = c(Perceptions, Priority, Policy, Health, Behavior),
    names_to  = "category",
    values_to = "flag"
  ) %>%
  # keep only “paper belongs to this category”
  filter(flag == 1) %>%
  dplyr::select(-flag)

>>>>

surveyCountries_long_clean$year <- ifelse(surveyCountries_long_clean$year == "2017 - 2021", 2019, surveyCountries_long_clean$year)
surveyCountries_long_clean$year <- ifelse(surveyCountries_long_clean$year == "2021 - 2022", 2021, surveyCountries_long_clean$year)

surveyCountries_long_clean_t <- surveyCountries_long_clean %>%
  # 1. build a single “years” string
  mutate(
    all_years = if_else(
      is.na(repeatedMeasure) | repeatedMeasure == "",
      as.character(year),
      paste(year, repeatedMeasure, sep = ";")
    ),
    # drop every space character
    all_years = str_remove_all(all_years, "\\s+")
  ) %>%
  separate_rows(all_years, sep = ";") %>%
  mutate(year = as.integer(all_years)) # %>%
  # dplyr::select(-all_years, -repeatedMeasure)

# 2) get total surveys by country
surveys_by_country <- surveyCountries_long_clean_t %>%
  filter(Inclusion == 1) %>% 
  group_by(iso3, .drop = F) %>% 
  count(iso3, name = "n_surveys") %>%
  arrange(desc(n_surveys))

@@

surveyCountries_long_clean_yearly <- surveyCountries_long_clean
surveyCountries_long_clean_yearly$year <- ifelse(surveyCountries_long_clean_yearly$year == "2017 - 2021", 2019, surveyCountries_long_clean_yearly$year)
surveyCountries_long_clean_yearly$year <- ifelse(surveyCountries_long_clean_yearly$year == "2021 - 2022", 2021, surveyCountries_long_clean_yearly$year)

surveyCountries_long_clean_yearly_t <- surveyCountries_long_clean_yearly %>%
  # 1. build a single “years” string
  mutate(
    all_years = if_else(
      is.na(repeatedMeasure) | repeatedMeasure == "",
      as.character(year),
      paste(year, repeatedMeasure, sep = ";")
    ),
    # drop every space character
    all_years = str_remove_all(all_years, "\\s+")
  ) %>%
  separate_rows(all_years, sep = ";") %>%
  mutate(year = as.integer(all_years)) %>%
  dplyr::select(-all_years, -repeatedMeasure)

df_cat_long <- surveyCountries_long_clean_yearly_t %>%
  pivot_longer(
    cols = c(Perceptions, Priority, Policy, Health, Behavior),
    names_to  = "category",
    values_to = "flag"
  ) %>%
  # keep only “paper belongs to this category”
  filter(flag == 1) %>%
  dplyr::select(-flag)

### 2025-12-05 draftFiguretwo_Ella.Rmd
fix bug year survey item

# bar plot with counts and percentage labels
pdf("plots/Figure2_category_paper_barplot.pdf")
ggplot(papers_long, aes(x = reorder(Topic, -n), y = n, fill = Topic)) +
  geom_col(color = "black") +
  geom_text(aes(label = percent(share, accuracy = 1)),
            vjust = -0.3, size = 5) +
  labs(x = NULL, y = "Number of papers") +
  scale_fill_manual(values = c(
    "Perceptions" = "#8B7B8B",
    "Behavior"    = "#EED8AE",
    "Priority"    = "#CDB5CD",
    "Policy"      = "#CD919E",
    "Health"      = "#8B475D",
    "None"        = "#F4A460")) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 30, hjust = 1, size = 15),
    axis.text.y = element_text(size = 15),
    axis.title = element_text(size = 16))
dev.off()

@@

surveys_long <- surveys %>%
  mutate(item_id = paste0(iso3, id)) %>% 
  distinct(item_id, category) %>%     # collapse multiple rows for same ID–category pair
  count(category, name = "n") %>% 
  mutate(share = n / length(unique(paste0(surveys$iso3, surveys$id)))) %>%
  rename(Topic = category)

# bar plot with counts and percentage labels
pdf("plots/Figure2_category_survey_barplot_share.pdf", width = 12, height = 10)
ggplot(surveys_long, aes(y = reorder(Topic, n), x = n, fill = Topic)) +
  geom_col(color = "black") +
  geom_text(aes(label = percent(share, accuracy = 1)),
            hjust = -0.3, size = 9) +
  labs(y = NULL, x = "Number of survey items covering each category") +
  scale_fill_manual(values = c(
    "Perceptions" = "#8B7B8B",
    "Behavior"    = "#EED8AE",
    "Priority"    = "#CDB5CD",
    "Policy"      = "#CD919E",
    "Health"      = "#8B475D",
    "None"        = "#F4A460")) +
  xlim(0, 1250) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.y = element_text(size = 22),
    axis.text.x = element_text(size = 22),
    axis.title = element_text(size = 23)
    )
dev.off()

@@

# 1. Prepare topic-level data (shares within income group, based on unique survey items)
surveys_topics <- surveys %>%
  filter(!is.na(Income.group)) %>%
  mutate(item_id = paste0(iso3, id)) %>%              # unique survey-level ID
  distinct(item_id, Income.group, category) %>%       # each item counts once per (Income.group, category)
  count(Income.group, category, name = "n_items") %>% # number of items mentioning category
  rename(Topic = category) %>%
  mutate(
    Income.group = factor(
      Income.group,
      levels = c("Low income", "Lower middle income", "Upper middle income", "High income")
    ),
    Topic = factor(
      Topic,
      levels = rev(c("Perceptions", "Behavior", "Priority", "Policy", "Health"))
    )
  )

# 2. Totals per income group (unique items)
surveys_totals <- surveys %>%
  filter(!is.na(Income.group)) %>%
  mutate(item_id = paste0(iso3, id)) %>%
  group_by(Income.group) %>%
  summarise(
    total = n_distinct(item_id),
    .groups = "drop"
  ) %>%
  mutate(
    Income.group = factor(
      Income.group,
      levels = c("Low income", "Lower middle income", "Upper middle income", "High income")
    )
  )

@@

# 1. Prepare topic-level data (shares within exposure group, based on unique survey items)
topics_data_survey_poll <- surveys %>%
  filter(!is.na(categoryPollution)) %>%
  mutate(item_id = paste0(iso3, id)) %>%                      # unique survey item ID
  distinct(item_id, categoryPollution, category) %>%          # each item counts once per (exposure, topic)
  count(categoryPollution, category, name = "n_items") %>%    # number of items mentioning category
  rename(Topic = category) %>%
  mutate(
    # optional: topic order, same as your income-group plot
    Topic = factor(
      Topic,
      levels = rev(c("Perceptions", "Behavior", "Priority", "Policy", "Health"))
    )
    # optional: control exposure group order, e.g.:
    # categoryPollution = factor(
    #   categoryPollution,
    #   levels = c("Air pollution", "Water pollution", "Other")  # adjust to your data
    # )
  )

# 2. Prepare totals row data (absolute counts of unique items per exposure group)
totals_data_survey_poll <- surveys %>%
  filter(!is.na(categoryPollution)) %>%
  mutate(item_id = paste0(iso3, id)) %>%
  group_by(categoryPollution) %>%
  summarise(
    total = n_distinct(item_id),
    .groups = "drop"
  )
  # optional: match factor ordering here too
  # %>%
  # mutate(
  #   categoryPollution = factor(
  #     categoryPollution,
  #     levels = c("Air pollution", "Water pollution", "Other")
  #   )
  # )

>>>>

# bar plot with counts and percentage labels
pdf("plots/Figure2_category_paper_barplot.pdf")
ggplot(papers_long, aes(x = reorder(Topic, -n), y = n, fill = Topic)) +
  geom_col(color = "black") +
  geom_text(aes(label = percent(share, accuracy = 1)),
            vjust = -0.3, size = 5) +
  labs(x = NULL, y = "Number of papers") +
  scale_fill_manual(values = c(
    "Perceptions" = "#8B7B8B",
    "Behavior"    = "#EED8AE",
    "Priority"    = "#CDB5CD",
    "Policy"      = "#CD919E",
    "Health"      = "#8B475D",
    "None"        = "#F4A460")) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 30, hjust = 1, size = 15),
    axis.text.y = element_text(size = 15),
    axis.title = element_text(size = 16)
    )
dev.off()

@@

surveys_long <- surveys %>%
  mutate(item_id = paste0(iso3, id, year)) %>% 
  distinct(item_id, category) %>%     # collapse multiple rows for same ID–category pair
  count(category, name = "n") %>% 
  mutate(share = n / length(unique(paste0(surveys$iso3, surveys$id)))) %>%
  rename(Topic = category)

# bar plot with counts and percentage labels
pdf("plots/Figure2_category_survey_barplot_share.pdf", width = 12, height = 10)
ggplot(surveys_long, aes(y = reorder(Topic, n), x = n, fill = Topic)) +
  geom_col(color = "black") +
  geom_text(aes(label = percent(share, accuracy = 1)),
            hjust = -0.3, size = 9) +
  labs(y = NULL, x = "Number of survey items covering each category") +
  scale_fill_manual(values = c(
    "Perceptions" = "#8B7B8B",
    "Behavior"    = "#EED8AE",
    "Priority"    = "#CDB5CD",
    "Policy"      = "#CD919E",
    "Health"      = "#8B475D",
    "None"        = "#F4A460")) +
  xlim(0, 1850) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.y = element_text(size = 22),
    axis.text.x = element_text(size = 22),
    axis.title = element_text(size = 23)
    )
dev.off()

@@

# 1. Prepare topic-level data (shares within income group, based on unique survey items)
surveys_topics <- surveys %>%
  filter(!is.na(Income.group)) %>%
  mutate(item_id = paste0(iso3, id, year)) %>%              # unique survey-level ID
  distinct(item_id, Income.group, category) %>%       # each item counts once per (Income.group, category)
  count(Income.group, category, name = "n_items") %>% # number of items mentioning category
  rename(Topic = category) %>%
  mutate(
    Income.group = factor(
      Income.group,
      levels = c("Low income", "Lower middle income", "Upper middle income", "High income")
    ),
    Topic = factor(
      Topic,
      levels = rev(c("Perceptions", "Behavior", "Priority", "Policy", "Health"))
    )
  )

# 2. Totals per income group (unique items)
surveys_totals <- surveys %>%
  filter(!is.na(Income.group)) %>%
  mutate(item_id = paste0(iso3, id, year)) %>%
  group_by(Income.group) %>%
  summarise(
    total = n_distinct(item_id),
    .groups = "drop"
  ) %>%
  mutate(
    Income.group = factor(
      Income.group,
      levels = c("Low income", "Lower middle income", "Upper middle income", "High income")
    )
  )

@@

# 1. Prepare topic-level data (shares within exposure group, based on unique survey items)
topics_data_survey_poll <- surveys %>%
  filter(!is.na(categoryPollution)) %>%
  mutate(item_id = paste0(iso3, id, year)) %>%                      # unique survey item ID
  distinct(item_id, categoryPollution, category) %>%          # each item counts once per (exposure, topic)
  count(categoryPollution, category, name = "n_items") %>%    # number of items mentioning category
  rename(Topic = category) %>%
  mutate(
    # optional: topic order, same as your income-group plot
    Topic = factor(
      Topic,
      levels = rev(c("Perceptions", "Behavior", "Priority", "Policy", "Health"))
    )
    # optional: control exposure group order, e.g.:
    # categoryPollution = factor(
    #   categoryPollution,
    #   levels = c("Air pollution", "Water pollution", "Other")  # adjust to your data
    # )
  )

# 2. Prepare totals row data (absolute counts of unique items per exposure group)
totals_data_survey_poll <- surveys %>%
  filter(!is.na(categoryPollution)) %>%
  mutate(item_id = paste0(iso3, id, year)) %>%
  group_by(categoryPollution) %>%
  summarise(
    total = n_distinct(item_id),
    .groups = "drop"
  )
  # optional: match factor ordering here too
  # %>%
  # mutate(
  #   categoryPollution = factor(
  #     categoryPollution,
  #     levels = c("Air pollution", "Water pollution", "Other")
  #   )
  # )
